Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections

### 2023-03-14 _paper.rmd
-Attempt to add tag for release. -Added Przemek.

---
title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
shorttitle: "Cheem, Exploring Local Explanations"
author:
- familyname: Spyrison
  othernames: Nicholas
  address: Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212
  email: nicholas.spyrison@monash.edu
  correspondingauthor: true
- familyname: Cook
  othernames: Dianne
  address: Monash University\newline ORCiD 0000-0002-3813-7155
  email: dicook@monash.edu
# - familyname: Marriott
#   othernames: Kim
#   address: Monash University\newline ORCiD 0000-0002-9813-0377
#   email: kim.marriott@monash.edu
year: 2022
date_submitted: "05/11"
abstract: "The increased predictive power of nonlinear models comes at the cost of interpretability of its terms. This trade-off has led to the emergence of eXplainable AI (XAI). XAI attempts to shed light on how models use predictors to arrive at a prediction with _local explanations_, a point estimate of the linear feature importance in the vicinity of one instance. These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface. Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output. The methods are implemented in the __R__ package __cheem__, available on CRAN."
keywords: "explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour"
bibliography: cheem_paper.bib
classoption: 11pt,twoside
output:
  bookdown::pdf_document2:
    template: jmlr_template.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    #citation_package: natbib
    latex_engine: pdflatex
---

>>>>

---
title: "Exploring Local Explanations of Nonlinear Models Using Animated Linear Projections"
shorttitle: "Cheem, Exploring Local Explanations"
author:
- familyname: Spyrison
  othernames: Nicholas
  address: Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212
  email: spyrison@gmail.com
  correspondingauthor: true
- familyname: Cook
  othernames: Dianne
  address: Monash University\newline ORCiD 0000-0002-3813-7155
  email: dicook@monash.edu
- familyname: Biecek
  othernames: Przemyslaw
  address:  Warsaw University of Technology and University of Warsaw\newline ORCiD 0000-0001-8423-1823
  email: przemyslaw.biecek@gmail.com
# - familyname: Marriott
#   othernames: Kim
#   address: Monash University\newline ORCiD 0000-0002-9813-0377
#   email: kim.marriott@monash.edu
year: 2022
date_submitted: "05/11"
abstract: "The increased predictive power of nonlinear models comes at the cost of interpretability of its terms. This trade-off has led to the emergence of eXplainable AI (XAI). XAI attempts to shed light on how models use predictors to arrive at a prediction with _local explanations_, a point estimate of the linear feature importance in the vicinity of one instance. These can be considered linear projections and can be further explored to understand better the interactions between features used to make predictions across the predictive model surface. Here we describe interactive linear interpolation used for exploration at any instance and illustrate with examples with categorical (penguin species, chocolate types) and quantitative (soccer/football salaries, house prices) output. The methods are implemented in the __R__ package __cheem__, available on CRAN."
keywords: "explainable artificial intelligence, nonlinear model interpretability, visual analytics, local explanations, grand tour, radial tour"
bibliography: cheem_paper.bib
classoption: 11pt,twoside
output:
  bookdown::pdf_document2:
    template: jmlr_template.tex
    fig_caption: yes
    fig_height: 5
    fig_width: 8
    keep_tex: yes
    number_sections: yes
    #citation_package: natbib
    latex_engine: pdflatex
---

### 2023-03-14 _paper.tex
JMLR and arXiv submission,
-Attempt to add tag for release. -Added Przemek.

\jmlrheading{}{2022}{}{MM/DD}{--}{}{%
Nicholas Spyrison and Dianne Cook}

@@

\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison and Cook}
\firstpageno{1}

@@

\author{%
\name Nicholas Spyrison \email \href{mailto:nicholas.spyrison@monash.edu}{\nolinkurl{nicholas.spyrison@monash.edu}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
%
\name Dianne Cook \email \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
}

@@

@@

>>>>

\jmlrheading{}{2022}{}{05/11}{--}{}{%
Nicholas Spyrison, Dianne Cook and Przemyslaw Biecek}

@@

\ShortHeadings{Cheem, Exploring Local Explanations}{Spyrison, Cook and Biecek}
\firstpageno{1}

@@

\author{%
\name Nicholas Spyrison \email \href{mailto:spyrison@gmail.com}{\nolinkurl{spyrison@gmail.com}} \\\addr Monash University\newline Australia\newline ORCiD 0000-0002-8417-0212 \AND
%
\name Dianne Cook \email \href{mailto:dicook@monash.edu}{\nolinkurl{dicook@monash.edu}} \\\addr Monash University\newline ORCiD 0000-0002-3813-7155 \AND
%
\name Przemyslaw Biecek \email \href{mailto:przemyslaw.biecek@gmail.com}{\nolinkurl{przemyslaw.biecek@gmail.com}} \\\addr Warsaw University of Technology and University of Warsaw\newline ORCiD 0000-0001-8423-1823 \AND
}

@@

\leavevmode\vadjust pre{\hypertarget{ref-de_cock_ames_2011}{}}%
De Cock, Dean. 2011. {``Ames, {Iowa}: {Alternative} to the {Boston} Housing Data as an End of Semester Regression Project.''} \emph{Journal of Statistics Education} 19 (3).

@@

\leavevmode\vadjust pre{\hypertarget{ref-prevek18_ames_2018}{}}%
prevek18. 2018. {``Ames {Housing} {Dataset}.''} \emph{Kaggle.com}. \url{https://kaggle.com/prevek18/ames-housing-dataset}.
